
## Problem 1

### Problem 1.a

P(clinton) = 4/7
P(sanders) = 3/7

P(state | clinton) = 1/4

P(voter | clinton) = 3/4

P(women | clinton) = 3/4

P(help | clinton) = 3/4

P(benghazi | clinton) = 0

P(state | sanders) = 1/3

P(voter | sanders) = 1/3

P(women | sanders) = 1/3

P(help | sanders) = 1/3

P(benghazi | sanders) = 1/3

P(clinton | "state voter women help benghazi") = 1/4 3/4 3/4 3/4 0 4/7 = 0

P(sanders | "state voter women help benghazi") = 1/3 1/3 1/3 1/3 1/3 3/7 = 1/1701

According to unsmoothed Naive Bayes, Bernie Sanders was more likely to have sent the email.
Since Clinton never used the word "Benghazi" in one of her emails, her posterior probability was zero.
If that word hadn't been considered, the algorithm would have determined that she was more likely to have sent the email.
The comparative rarity of each individual word makes it very likely that the posterior probability will be zero.

### 1.b

P(clinton) = 5/8
P(sanders) = 1/2

P(state | clinton) = 2/5

P(voter | clinton) = 4/5

P(women | clinton) = 4/5

P(help | clinton) = 4/5

P(benghazi | clinton) = 1/5

P(state | sanders) = 1/2

P(voter | sanders) = 1/2

P(women | sanders) = 1/2

P(help | sanders) = 1/2

P(benghazi | sanders) = 1/2

P(clinton | "state voter women help benghazi") = 2/5 4/5 4/5 4/5 1/5 5/8 = 640/25000 = .0256

P(sanders | "state voter women help benghazi") = 1/2 1/2 1/2 1/2 1/2 1/2 = 1/64 = 0.015625

After Laplace smoothing is added, the algorithm indicates that Clinton is more likely to have written the email.

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
